---
title: "Maryland Poison Center Graphics"
author: "Leah Jager"
date: "6/26/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

My goal is to make a map of Maryland by county that is colored by some count or rate from the MPC data set I have created.  I would like to make this map for each year, 2006-2018, and then animate it to show changes over time.

```{r}
library(tidyverse)
library(ggmap)
library(maps)
```

First, read in the MPC data:
```{r}
mpcData <- read_csv("MPCdataFINAL.csv")
names(mpcData)
dim(mpcData)
glimpse(mpcData)
```

Now just playing with making maps of Maryland:
```{r}
#making a plot of counties in maryland using the plotting within map package
map('county', 'maryland', fill = TRUE, col = palette())

#making a plot of counties in maryland using ggplot
md <- map_data('county', 'maryland')
dim(md)
head(md)
tail(md)

# fill is black and county lines are in white
ggplot() + 
  geom_polygon(data = md, aes(x=long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
  theme_bw()

# no fill, red outline
ggplot() + 
  geom_polygon(data = md, aes(x=long, y = lat, group = group), fill="NA", color="red") + 
  coord_fixed(1.3) + 
  theme_bw()

# orange fill, black outline, go orioles?
ggplot() + 
  geom_polygon(data = md, aes(x=long, y = lat, group = group), fill="orange", color="black") + 
  coord_fixed(1.3) + 
  theme_bw()

# give each county a different color
ggplot() + 
  geom_polygon(data = md, aes(x=long, y = lat, fill=subregion, group = group), color="white") + 
  coord_fixed(1.3) + 
  theme_bw()

# give each county a different color and remove color legend
ggplot() + 
  geom_polygon(data = md, aes(x=long, y = lat, fill=subregion, group = group), color="white") + 
  coord_fixed(1.3) + 
  theme_bw() + 
  guides(fill=FALSE) 

# get rid of gridlines
ggplot() + 
  geom_polygon(data = md, aes(x=long, y = lat, fill=subregion, group = group), color="white") + 
  coord_fixed(1.3) + 
  theme_nothing() + 
  guides(fill=FALSE) 

# add state outline in black
state.md <- map_data('state', 'maryland')
ggplot() + 
  geom_polygon(data = md, aes(x=long, y = lat, fill=subregion, group = group), color="white") + 
  geom_polygon(data = state.md, aes(x=long, y=lat, group=group), color="black", fill=NA) +
  coord_fixed(1.3) + 
  theme_nothing() + 
  guides(fill=FALSE) 
```

Now going to try to color by the number of human exposures:
```{r}
mpc2018 <- mpcData %>%
  filter(Year==2018) %>%
  mutate(subregion=tolower(County))

mpc2018$subregion
unique(md$subregion)

# need to remove the , md and the apostrophes
mpc2018$subregion = str_replace_all(mpc2018$subregion,", md", "")
mpc2018$subregion = str_replace_all(mpc2018$subregion," county", "")
mpc2018$subregion = str_replace_all(mpc2018$subregion,"’", "")
mpc2018$subregion = str_replace_all(mpc2018$subregion,"[.]", "")

mpc2018$subregion
unique(md$subregion)

plotData <- inner_join(md, mpc2018, by="subregion")

dim(plotData)
dim(table(plotData$subregion))

ggplot() + 
  geom_polygon(data = plotData, aes(x=long, y = lat, fill=TotalHumanExposures, group = group), color="white") + 
  geom_polygon(data = state.md, aes(x=long, y=lat, group=group), color="black", fill=NA) +
  coord_fixed(1.3) + 
  theme_bw()

```

These plots are pretty meaningless without turning them into a rate using the population, but first lets just make a fdw more:
```{r}
ggplot() + 
  geom_polygon(data = plotData, aes(x=long, y = lat, fill=MedOutcome_Death, group = group), color="white") + 
  geom_polygon(data = state.md, aes(x=long, y=lat, group=group), color="black", fill=NA) +
  coord_fixed(1.3) + 
  theme_bw()

ggplot() + 
  geom_polygon(data = plotData, aes(x=long, y = lat, fill=Intentional_SuspectedSuicide, group = group), color="white") + 
  geom_polygon(data = state.md, aes(x=long, y=lat, group=group), color="black", fill=NA) +
  coord_fixed(1.3) + 
  theme_bw()

ggplot() + 
  geom_polygon(data = plotData, aes(x=long, y = lat, fill=Age_Under12mo, group = group), color="white") + 
  geom_polygon(data = state.md, aes(x=long, y=lat, group=group), color="black", fill=NA) +
  coord_fixed(1.3) + 
  theme_bw()
```

First we can consider a scale change:
```{r}
ggplot() + 
  geom_polygon(data = plotData, aes(x=long, y = lat, fill=Age_Under12mo, group = group), color="white") + 
  geom_polygon(data = state.md, aes(x=long, y=lat, group=group), color="black", fill=NA) +
  coord_fixed(1.3) + 
  theme_bw() + scale_fill_gradient(trans="log10")
```

Now I need to find some population data by county:
```{r}
popData <- read_csv("./PEP_2018_PEPANNRES/PEP_2018_PEPANNRES.csv")
glimpse(popData)

pop2018 <- popData %>%
  select(respop72018, `GEO.display-label`) %>%
  mutate(subregion=tolower(`GEO.display-label`))

unique(plotData$subregion)
pop2018$subregion

# need to remove the , md and the apostrophes
pop2018$subregion = str_replace_all(pop2018$subregion,", maryland", "")
pop2018$subregion = str_replace_all(pop2018$subregion," county", "")
pop2018$subregion = str_replace_all(pop2018$subregion,"'", "")
pop2018$subregion = str_replace_all(pop2018$subregion,"[.]", "")

unique(plotData$subregion)
pop2018$subregion

plotData <- inner_join(plotData, pop2018, by="subregion")
plotData <- plotData %>%
  mutate(THEperCap=TotalHumanExposures/respop72018)

ggplot() + 
  geom_polygon(data = plotData, aes(x=long, y = lat, fill=THEperCap, group = group), color="white") + 
  geom_polygon(data = state.md, aes(x=long, y=lat, group=group), color="black", fill=NA) +
  coord_fixed(1.3) + 
  theme_bw()

```